Compositional equivalences based on Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2019-11-13 2016-Weak-FH-Bisim/WeakBisim.tex
More on Figure 2...
Figs de l'annexe D
traces execution

Fig. \ref{SimpleProt:Spec} (right) shows a graphical view of the pNet of the system \emph{specification}:
the full system should be equivalent (e.g. through weak bisimulation)
to the two processes connected simply through a perfect medium. Fig. \ref{SimpleProt:Spec}, right, shows the pNet representing this global specification.

@@

\begin{figure}[t]
  \centerline{\includegraphics[width=12cm]{XFIG/SimpleProt2-pNet}}
  \caption{Composed pNet with the Simple Protocol Implementation}  \label{SimpleProt:Impl}
\end{figure}

@@

We take in this article a semantics and logical understanding of these
automata. Open automata are closed by a simple form of refinement that
allows us to refine the predicate, or substitute any free variable by
an expression. More formally, let $\Post$ be any substitution and
$\Pred$ any predicate suppose $\vars(t)\cap \dom(\Post)=\emptyset$ and
$\vars(t')\cap \dom(\Post)=\emptyset$. Then we have the following
implication:

@@

The equivalence we need is a strong bisimulation between
pNets having exactly the same Holes with the same sorts, but using a
flexible matching 
between open transition, to accommodate comparisons between pNet
expressions with different architectures.

We define now a bisimulaiton relation adapted to open automata and their parametric nature. The relation relates states of the open automaton and states equivalence between the open transitions between the states. Its key characteristics are 1) the introduction of predicates in the bisimulation relation: as states may contain variables, relation between states may depend on the value of the variables; 2) the bisimulation property relates elements of the open transitions and take into account predicates over variables, actions of the holes, and state modifications.
 We name it FH-bisimulation,
 as a short cut for the ``Formal Hypotheses'' manipulated in the
 transitions, but also as a reference to the work of De Simone~\cite{deSimone85},
 that pioneered this idea.

Let $\mathcal{R}=\{(s,t|\Pred_{s,t})\}$ be a relation over the set $\mathcal{S}_1$ and 
$\mathcal{S}_2$ constrained by a predicate
More precisely, for any pair $(s,t)$, there is a 
   single
      $(s,t|\Pred_{s,t})\in\mathcal{R}$  stating that $s$ and $t$ are related 
      if $\Pred_{s,t}$       is 
      t\\True, i.e. the states are related when the variables in $s$ and $t$ verify the 
      predicate $\Pred_{s,t}$.
 FH-bisimulation is defined formally: 
 \begin{definition}[Strong FH-bisimulation]\label{def-FH-bisim} ~\\

@@

\begin{figure}[t]
   \includegraphics[width=5.5cm]{XFIG/SimpleProt-Schema}
   \includegraphics[width=6.5cm]{XFIG/SimpleProt-Spec}
   \caption{Specification pNet Schema and Specification}
   \label{App:SimpleProt:Spec}

@@

\begin{figure}[t]
  \centerline{\includegraphics[width=12cm]{XFIG/SimpleProt-pNet}}
  \caption{Composed pNet with the Simple Protocol Implementation}  \label{App:SimpleProt:Impl}
\end{figure}

@@

The corresponding generated Open Automaton was given in Figure \ref{},
page \pageref{}.

@@

\begin{lstlisting}[basicstyle=\scriptsize\ttfamily, language=java, frame=single]
SimpleProtImpl2
--------------------
SV6 SVs: <p_send(m), in(m), _>-->in(m)
SV7 SVs: <p_a, _, _>-->p_a, p_a!=(p_send(m))
SV8 SVs: <_, out(m, c), q_recv(m, c)>-->out(m, c)
SV9 SVs: <_, _, q_a>-->q_a, q_a!=(q_recv(m, c))
SV11 SVs: <_, _tau_, _>-->_tau_
SV12 SVs: <_, _tau_, _>-->_tau_
SV13 SVs: <_, _tau_, _>-->_tau_
SV14 SVs: <_, _tau_, _>-->_tau_
SV15 SVs: <_, _tau_, _>-->_tau_
Holes: P
Subnets: medium
- medium
- SV0 SVs: <in(m), _, _>-->in(m)
- SV1 SVs: <s_send(m, c), m_recv(m, c), _>-->_tau_
- SV2 SVs: <_, m_send(m), r_recv(m)>-->_tau_
- SV3 SVs: <s_ack, _, r_ack>-->_tau_
- SV4 SVs: <s_error, m_error, _>-->_tau_
- SV5 SVs: <_, _, out(m, c)>-->out(m, c)
- SV10 SVs: <_, _tau_, _>-->_tau_
- pLTS: Sender
- - Transition: s0---in(msg)-->s1, { s1_msg := msg s1_c := 0}
- - Transition: s1---s_send(s1_msg, s1_c)-->s2, {}
- - Transition: s2---s_ack-->s0, {}
- - Transition: s2---s_error-->s1, { s1_c := s2_c+1}
- - Variable: s1_c, State: s1, Value(s) = 0,s2_c+1,
- - Variable: s1_msg, State: s1, Value(s) = msg,
- pLTS: M1
- - Transition: a0---m_recv(msg, c)-->a1, { a1_msg := msg a1_c := c}
- - Transition: a1---m_send(a1_msg, a1_c)-->a0, {}
- - Transition: a1---_tau_-->a2, {}
- - Transition: a2---m_error-->a0, {}
- - Variable: a1_msg, State: a1, Value(s) = msg,
- - Variable: a1_c, State: a1, Value(s) = c,
- pLTS: Receiver
- - Transition: r0---r_recv(msg, c)-->r1, { r1_msg := msg r1_c := c}
- - Transition: r1---out(r1_msg, r1_c)-->r2, {}
- - Transition: r2---r_ack-->r0, {}
- - Variable: r1_msg, State: r1, Value(s) = msg,
- - Variable: r1_c, State: r1, Value(s) = c,
Holes: Q
\end{lstlisting}

Finaly, we list here, in raw format, the 24 OTs computed by the tool:

\begin{lstlisting}[basicstyle=\scriptsize\ttfamily, language=java, frame=single]
INFO  SimpleProtImpl2 - Total: 24 OTs
INFO  SimpleProtImpl2 - 
		{P|->p_send(m:sva_SV6:1:1)},   [((in(msg))=:ra:Sender:121:1)/\(:ra:Sender:121:1=(in(m:sva_SV0:12:1)))/\((in(m:sva_SV0:12:1))=(in(m:sva_SV6:1:1)))],   {s1_msg := msg, s1_c := 0}
	OT:1 ------------------------------------------------------------
		<s0_a0_r0>-----in(m:sva_SV6:1:1)----><s1_a0_r0>

@@

INFO  SimpleProtImpl2 - 
		{},   [((s_send(s1_msg, s1_c))=:ra:Sender:1144:1)/\(:ra:Sender:1144:1=(s_send(m:sva_SV1:114:2, c:sva_SV1:114:2)))/\((m_recv(msg, c))=:ra:M1:1145:1)/\(:ra:M1:1145:1=(m_recv(m:sva_SV1:114:2, c:sva_SV1:114:2)))/\(_tau_=_tau_)],   {a1_msg := msg, a1_c := c}
	OT:6 ------------------------------------------------------------
		<s1_a0_r0>-----_tau_----><s2_a1_r0>

INFO  SimpleProtImpl2 - 
		{},   [((s_send(s1_msg, s1_c))=:ra:Sender:1174:1)/\(:ra:Sender:1174:1=(s_send(m:sva_SV1:117:2, c:sva_SV1:117:2)))/\((m_recv(msg, c))=:ra:M1:1175:1)/\(:ra:M1:1175:1=(m_recv(m:sva_SV1:117:2, c:sva_SV1:117:2)))/\(_tau_=_tau_)],   {a1_msg := msg, a1_c := c}
	OT:7 ------------------------------------------------------------
		<s1_a0_r0>-----_tau_----><s2_a1_r0>

INFO  SimpleProtImpl2 - 
		{},   [((s_send(s1_msg, s1_c))=:ra:Sender:1204:1)/\(:ra:Sender:1204:1=(s_send(m:sva_SV1:120:2, c:sva_SV1:120:2)))/\((m_recv(msg, c))=:ra:M1:1205:1)/\(:ra:M1:1205:1=(m_recv(m:sva_SV1:120:2, c:sva_SV1:120:2)))/\(_tau_=_tau_)],   {a1_msg := msg, a1_c := c}
	OT:8 ------------------------------------------------------------
		<s1_a0_r0>-----_tau_----><s2_a1_r0>

INFO  SimpleProtImpl2 - 
		{},   [((s_send(s1_msg, s1_c))=:ra:Sender:1234:1)/\(:ra:Sender:1234:1=(s_send(m:sva_SV1:123:2, c:sva_SV1:123:2)))/\((m_recv(msg, c))=:ra:M1:1235:1)/\(:ra:M1:1235:1=(m_recv(m:sva_SV1:123:2, c:sva_SV1:123:2)))/\(_tau_=_tau_)],   {a1_msg := msg, a1_c := c}
	OT:9 ------------------------------------------------------------
		<s1_a0_r0>-----_tau_----><s2_a1_r0>

INFO  SimpleProtImpl2 - 
		{},   [((s_send(s1_msg, s1_c))=:ra:Sender:1264:1)/\(:ra:Sender:1264:1=(s_send(m:sva_SV1:126:2, c:sva_SV1:126:2)))/\((m_recv(msg, c))=:ra:M1:1265:1)/\(:ra:M1:1265:1=(m_recv(m:sva_SV1:126:2, c:sva_SV1:126:2)))/\(_tau_=_tau_)],   {a1_msg := msg, a1_c := c}
	OT:10 ------------------------------------------------------------
		<s1_a0_r0>-----_tau_----><s2_a1_r0>

INFO  SimpleProtImpl2 - 
		{P|->p_a:sva_SV7:1:2},   [(forall:m:sva_SV7:1:2. (p_a:sva_SV7:1:2!=(p_send(m:sva_SV7:1:2)))],   {}
	OT:11 ------------------------------------------------------------
		<s2_a1_r0>-----p_a:sva_SV7:1:2----><s2_a1_r0>

INFO  SimpleProtImpl2 - 
		{Q|->q_a:sva_SV9:1:4},   [(forall:m:sva_SV9:1:4,c:sva_SV9:1:4. (q_a:sva_SV9:1:4!=(q_recv(m:sva_SV9:1:4, c:sva_SV9:1:4)))],   {}
	OT:12 ------------------------------------------------------------
		<s2_a1_r0>-----q_a:sva_SV9:1:4----><s2_a1_r0>

INFO  SimpleProtImpl2 - 
		{},   [(_tau_=:ra:M1:11420:1)/\(:ra:M1:11420:1=_tau_)/\(_tau_=_tau_)],   {}
	OT:13 ------------------------------------------------------------
		<s2_a1_r0>-----_tau_----><s2_a2_r0>

INFO  SimpleProtImpl2 - 
		{},   [(_tau_=:ra:M1:11720:1)/\(:ra:M1:11720:1=_tau_)/\(_tau_=_tau_)],   {}
	OT:14 ------------------------------------------------------------
		<s2_a1_r0>-----_tau_----><s2_a2_r0>

INFO  SimpleProtImpl2 - 
		{},   [(_tau_=:ra:M1:12020:1)/\(:ra:M1:12020:1=_tau_)/\(_tau_=_tau_)],   {}
	OT:15 ------------------------------------------------------------
		<s2_a1_r0>-----_tau_----><s2_a2_r0>

INFO  SimpleProtImpl2 - 
		{},   [(_tau_=:ra:M1:12320:1)/\(:ra:M1:12320:1=_tau_)/\(_tau_=_tau_)],   {}
	OT:16 ------------------------------------------------------------
		<s2_a1_r0>-----_tau_----><s2_a2_r0>

INFO  SimpleProtImpl2 - 
		{},   [(_tau_=:ra:M1:12620:1)/\(:ra:M1:12620:1=_tau_)/\(_tau_=_tau_)],   {}
	OT:17 ------------------------------------------------------------
		<s2_a1_r0>-----_tau_----><s2_a2_r0>

INFO  SimpleProtImpl2 - 
		{P|->p_a:sva_SV7:1:2},   [(forall:m:sva_SV7:1:2. (p_a:sva_SV7:1:2!=(p_send(m:sva_SV7:1:2)))],   {}
	OT:18 ------------------------------------------------------------
		<s2_a2_r0>-----p_a:sva_SV7:1:2----><s2_a2_r0>

INFO  SimpleProtImpl2 - 
		{Q|->q_a:sva_SV9:1:4},   [(forall:m:sva_SV9:1:4,c:sva_SV9:1:4. (q_a:sva_SV9:1:4!=(q_recv(m:sva_SV9:1:4, c:sva_SV9:1:4)))],   {}
	OT:19 ------------------------------------------------------------
		<s2_a2_r0>-----q_a:sva_SV9:1:4----><s2_a2_r0>

INFO  SimpleProtImpl2 - 
		{},   [(s_error=:ra:Sender:11413:1)/\(:ra:Sender:11413:1=s_error)/\(m_error=:ra:M1:11414:1)/\(:ra:M1:11414:1=m_error)/\(_tau_=_tau_)],   {s1_c := s2_c+1}
	OT:20 ------------------------------------------------------------
		<s2_a2_r0>-----_tau_----><s1_a0_r0>

INFO  SimpleProtImpl2 - 
		{},   [(s_error=:ra:Sender:11713:1)/\(:ra:Sender:11713:1=s_error)/\(m_error=:ra:M1:11714:1)/\(:ra:M1:11714:1=m_error)/\(_tau_=_tau_)],   {s1_c := s2_c+1}
	OT:21 ------------------------------------------------------------
		<s2_a2_r0>-----_tau_----><s1_a0_r0>

INFO  SimpleProtImpl2 - 
		{},   [(s_error=:ra:Sender:12013:1)/\(:ra:Sender:12013:1=s_error)/\(m_error=:ra:M1:12014:1)/\(:ra:M1:12014:1=m_error)/\(_tau_=_tau_)],   {s1_c := s2_c+1}
	OT:22 ------------------------------------------------------------
		<s2_a2_r0>-----_tau_----><s1_a0_r0>

INFO  SimpleProtImpl2 - 
		{},   [(s_error=:ra:Sender:12313:1)/\(:ra:Sender:12313:1=s_error)/\(m_error=:ra:M1:12314:1)/\(:ra:M1:12314:1=m_error)/\(_tau_=_tau_)],   {s1_c := s2_c+1}
	OT:23 ------------------------------------------------------------
		<s2_a2_r0>-----_tau_----><s1_a0_r0>

INFO  SimpleProtImpl2 - 
		{},   [(s_error=:ra:Sender:12613:1)/\(:ra:Sender:12613:1=s_error)/\(m_error=:ra:M1:12614:1)/\(:ra:M1:12614:1=m_error)/\(_tau_=_tau_)],   {s1_c := s2_c+1}
	OT:24 ------------------------------------------------------------
		<s2_a2_r0>-----_tau_----><s1_a0_r0>

INFO  SimpleProtImpl2 - Total: 24 SATISFIABLE OTs
INFO  SimpleProtImpl2 -        278 UNSATISFIABLE OTs
INFO  SimpleProtImpl2 - Total Time: 1056 ms

@@

\newpage
  \section{Running example: Hierarchical Broadcast}

Hierarchical Broadcast (HB) is a way of implementing broadcasting
algorithms within a large set of processes, with better performances
and scalability. There have been many proposals for hierarchical
broadcast algorithms (cite...), most of them focussing on specific
architectures or features (in particular fault tolerance), and on
performance analysis. Very few have given a formal description of the
algorithm, allowing to reason about their safety properties.
In this paper, we choose HB as a running example because it is a
complex interaction pattern on which the topology and the
parameterised nature plays a crucial role.

We give here the description of one simple HB algorithm, inspired
from\cite{Taguchi03}. For the formal description, we define a
pNet template, that can be used as a building node in a hierarchical
assembly of processes, organised in a tree of sub-groups. Inside each
sub-group, messages are sent using  local broadcast; associated to
each sub-group is a ``gateway'' in charge of the
transmission of messages, up and down, between the subgroups. The
gateways are described as generic pLTSs, playing the role of
``controllers'' in the pNet node.

\begin{figure}[t]
\centerline{  \includegraphics[width=0.9\linewidth]{XFIG/SchemaHB2}}
  \caption{Principle of a Hierarchical Broadcast structure and protocol}
  \label{fig:hb-scenario}
\end{figure}

Figure \ref{fig:hb-scenario} shows an instantiation of a HB
structure, and a possible scenario:
\begin{itemize}
\item
  The HB tree has 3 sub-groups, each with its gateway process, and 6
  processes scattered in the sub-groups.
\item
  The path in green shows the sequence of message exchanges when Process P21 (within
  group G2) sends a message to process P111 (within group G11).
\item
  As the target is not within the local group to which P21 belongs,
  the initial message $m(111)$ is sent up to group 2, and
  arrives in gateway $gw2$.
\item
  Now the destination belongs to a group that is a sibling of G2, so
  it is broadcast to all its siblings (sent as $bm(111)$, and
  received as $m(111)$). It will be simply discarded by
  gateways or processes that are not involved (here P3), while $gw1$
  will recognize as being of interest, and in next step send it down
  to its group.
\item
  Sending down is again a broadcast, to all members of sub-group
  G1. Process P12 will ignore it, while gateway $gw11$ in turn
  will recognize and forward the message.
\item
  Within group G11, process P111 will finally receives message
  $m(111)$, that terminates our scenario.
\end{itemize}

Along this paper, we will use this example to illustrate the use of
pNets as a structured semantic model, with a stress on the fact that
we represent here HB as a program schema, with unspecified process
arguments, using open pNets. We will show how the symbolic operational
semantics of open pNets computes the dynamics of the HB in term of a
symbolic representation that we call open-transitions, and how we can
prove equivalences between open systems.

\begin{example} \emph{First step: define the action algebra, with data domains,
    operators and predicates.}
  For our HB structure, we use send and receive messages encoded
  respectively as $m(arg)$ and $im(?var)$ with
  $arg$ of type \texttt{Address}. HB has a message-based
  structure, and by convention action names for action reception will
  be prefix by the letter "i". We abstract away
  from the payload of the messages, keeping only target addresses in the HB
  tree as an argument to all messages. Addresses are finite sequences
  of natural numbers written \texttt{s.i}
  with \texttt{s} a sequence and i a number. In concrete instantiations, we
  assume numbers are simple digits, and we write the sequences without
  the dot, as in ``121''. Then in the model of the
  gateways, we must be able to decide whether an address is below/above the
  current address, or in some specific relation, using
  predicates:\\
  - \emph{i==j} Syntactic equality between two sequences.\\
  - \emph{is\_ancestor(i,j)}: $i$ is a (strict) prefix of $j$.\\
  - \emph{is\_direct\_relative(i,j)}: $i$ and $j$ have a common prefix $k$, $i=k.i1$, $j=k.j1$, with $i1$ of length 1, but $j1$ can be of length equal or greater than 1 ($j$ is a brother or a nephew in the genealogy).\\
  - \emph{is\_other\_relative(i,j)}: all other cases.\\
  Note that these 4 predicates are exclusive.
\end{example}

\begin{example}  \emph{Second step: define the gateway behaviour as a pLTS.}

\begin{figure}[h]
\centerline{  \includegraphics[width=0.85\linewidth]{ATG/GatewayHB}}
\caption{Behaviour of a Gateway controller %\TODO{ERIC: change the action syntax}
  }
  \label{fig:hb-HBgateway}
\end{figure}
Figure \ref{fig:hb-HBgateway} shows the pLTS representing the generic
gateway controller. From its initial state, it can receive messages
$ium(?tgt)$ from the subgroup below, or $im(?tgt)$
coming either from above or from an horizontal broadcast. The details will
be explained later, together with the model structure and the
protocol.
To simplify the notation in the pLTS drawings, we use the following
conventions: state variables are implicitely subscripted by the
adequate state number (e.g. $ium(?tgt)$ means $ium(?tgt_2)$), and we
have ``global'' variables that are defined in all states (here $gid$
means $gid_0,gid_1,gid_2$).

Note that we use \texttt{if-then-else} or \texttt{case}
constructs in the pLTS labels. These are mere abbreviations, that can
be expanded as a set of transitions carrying simple labels with adequate guards.
\end{example}

\begin{example} \emph{Sorts of pLTSs:}
  The sort of the gateway in Figure \ref{fig:hb-HBgateway} is:\\
  $\large\{ um(tgt_s), im(tgt_s), dm(tgt_s), bm(tgt_s), um(tgt_s),
  ium(tgt_s), DISCARD, \\ERROR(name) \large\}_{s\in\{1,2\}}$
\end{example}

\begin{figure}[t]
  \includegraphics[width=1.0\linewidth]{ATG/Inst5}
  \caption{A Hierarchical Broadcast pNet Structure}\label{fig:hb-pnet-graphical}
 \end{figure}

\begin{example} \emph{pNet structure, and graphical representation}  
Here we show some parts of the pNet tree expressing our HB
structure. Figure \ref{fig:hb-pnet-graphical} shows the full pNet
model in a graphical format, where each box represents a pNet node
(Groot, G1, ..., G11), a pLTS (gw1, ..., gw11), or a process (P3, ...,
P111).
The actions in the sorts of the nodes appear as ports on the boxes,
and the synchronisation vectors as arrows, or multipoint ``webs''
drawn as ellipses. For
example, the webs
%labelled \texttt{tau}
in the Groot box
  correspond to synchronisation vectors expressing a broadcast
  communication from one of the subnets of Groot (gw1, gw2, or P3), to
  the others.

  Formally, the root pNet has a structure $<gw1, gw2, G1, G2, P3>$,
  with G1 and G2 sub-pNets, their associated gateways gw1, gw2
  pLTSs, and P3 a process Hole. We have 4 possible broadcast communications at the
  root, corresponding to as many synchronisation vectors; e.g. if gw2 is the sender:
  $< im(a), bm(a), -, -, im(a)> \to\!\underline{bm(a)}$.

  We define the synchronisation vectors in a parameterised way
  (independently of a specific HB network configuration) for any
  pNet node in a HB structure. Let $\widetilde{i}$ be the position of this pNet in the
  structure.
   We define here 4 of these synchronisation vectors, that we will
   use in the next sections to illustrate the semantic rules.
   %; the full table is listed in Appendix \ref{Appendix:SynchVectors}.
   In the
  vectors below,
  $\widetilde{I_G}=\{gw\widetilde{i}1,..,gw\widetilde{i}p,G\widetilde{i}1,..,G\widetilde{i}p\}$
   and
  $\widetilde{I_P}=\{P\widetilde{i}1,..,P\widetilde{i}p\}$ are the sets
  containing respectively the indexes of sub-nets and holes of the
  pNet node, $\widetilde{I_{G+P}}$ their union.

  \begin{figure}[h!]
  \label{fig:HBvectors}
  \begin{tabular}{|p{1.4cm}|p{11cm}|}
%% \hline\hline
%% \multicolumn{3}{|l|}{Hierarchical broadcasting}\\
  \hline
  \hline
$(UG)_{\widetilde{i}k}$\newline
 $k\leq g$&\raisebox{-2pt}{
 $\begin{array}[t]{@{}c@{}c@{}c@{}c@{}c@{}c@{}c@{}c@{}}
 &gw\widetilde{i}1,..&,gw\widetilde{i}k,&..,gw\widetilde{i}g,G\widetilde{i}1,..&,G\widetilde{i}k,&..,G\widetilde{i}g,P\widetilde{i}1,..,P\widetilde{i}p\\
<\!& -\,-\,-\    &  ,ium(\widetilde{t}),& -\ -\ -\  &,um(\widetilde{t}),&, -\ -\ -\
&\!>\longrightarrow \underline{um_{\widetilde{i}.k}(\widetilde{t})}
% $
%     \symb{sv}_{u(\widetilde{i}k)}\triangleq
%     <ium_{\widetilde{i}.k}(\widetilde{t})\shortotimes
%     -^{\widetilde{I_P}}\shortotimes
%     um_{\widetilde{i}.k}(\widetilde{t})\shortotimes
%  -^{{\widetilde{I_G}\backslash\{\widetilde{i}\cdot
%         k\}}}>\longrightarrow
%     \underline{um(\widetilde{t})}$
\end{array}$}  \raisebox{-19pt}{~}  \\
\hline
     &{\ERIC{Up gateway}: message m is sent
  up from a group to its gateway, received by the ium action. The
  resulting action is visible as the (synchronised) $\underline{um}$ }\\
\hline
 $(UM)_{\widetilde{i}k}$\newline$k\leq g$
 &\raisebox{-2pt}{$\begin{array}[t]{@{}c@{}c@{}c@{}c@{}c@{}c@{}c@{}c@{}}
 &gw\widetilde{i}1,..&,gw\widetilde{i}k,&..,gw\widetilde{i}g,G\widetilde{i}1,..,G\widetilde{i}g,P\widetilde{i}1,..,P\widetilde{i}p\\
 <\!& -\,-\,-\    &  ,um(\widetilde{t}),& -\ -\ -\   -\ -\ -\
 &\!>\longrightarrow \underline{um(\widetilde{t})}
 \end{array}$}\raisebox{-18pt}{~}\\
\hline
$(UM)_{\widetilde{i}k}$\newline$k\leq p$
&\raisebox{-2pt}{$\begin{array}[t]{@{}c@{}c@{}c@{}c@{}c@{}c@{}c@{}c@{}}
&gw\widetilde{i}1,..,gw\widetilde{i}g,G\widetilde{i}1,..,,G\widetilde{i}g,P\widetilde{i}1,..&,P\widetilde{i}k,&..,P\widetilde{i}p\\
<\!& -\ -\ -\      -\ -\ -\ &,um(\widetilde{t}),&   -\, -\, -
&\!>\longrightarrow \underline{um(\widetilde{t})}
\end{array}$}\raisebox{-18pt}{~}
 % &
%$\symb{sv}_{!\widetilde{i}}\triangleq
%    <um_{\widetilde{i}.k}(\widetilde{t})\shortotimes
%    -^{\widetilde{I_{G+P}}\backslash\{\widetilde{i}\cdot k\}}\shortotimes
%    -^{\widetilde{I_G}}>\longrightarrow
%    um_{\widetilde{i}.k}(\widetilde{t})$
\\
\hline
     &{\ERIC{Up Message}: message m is sent up from a process
  or a gateway to the enclosing (= parent) group}\\
\hline

>>>>

Fig. \ref{SimpleProt:Spec} (right) shows a graphical view of the pNet of the system \emph{specification}:
the full system should be equivalent (e.g. through weak bisimulation)
to the two processes connected simply through a perfect medium.

@@

\begin{figure}[t]
  \centerline{\includegraphics[width=12cm]{XFIG/SimpleProt2-pNet-tau}}
  \caption{Composed pNet with the Simple Protocol Implementation}  \label{SimpleProt:Impl}
\end{figure}

@@

We take in this article a semantics and logical understanding of these
automata. Open automata are closed by a simple form of refinement that
allows us to refine the predicate, or substitute any free variable by
an expression. More formally, let $\Post$ be any substitution and
$\Pred$ any predicate; suppose $\vars(t)\cap \dom(\Post)=\emptyset$ and
$\vars(t')\cap \dom(\Post)=\emptyset$. Then we have the following
implication:

@@

The equivalence we need is a strong bisimulation between
pNets having exactly the same Holes with the same sorts, but using a
flexible matching 
between open transitions, to accommodate comparisons between pNet
expressions with different architectures.

We define now a bisimulation relation adapted to open automata and their parametric nature. The relation relates states of the open automaton and states equivalence between the open transitions between the states. Its key characteristics are 1) the introduction of predicates in the bisimulation relation: as states may contain variables, relation between states may depend on the value of the variables; 2) the bisimulation property relates elements of the open transitions and take into account predicates over variables, actions of the holes, and state modifications.
 We name it FH-bisimulation,
 as a short cut for the ``Formal Hypotheses'' over the holes behavior manipulated in the
 transitions, but also as a reference to the work of De Simone~\cite{deSimone85},
 that pioneered this idea.

Let $\mathcal{R}=\{(s,t|\Pred_{s,t})\}$ be a relation over the set $\mathcal{S}_1$ and 
$\mathcal{S}_2$ constrained by a predicate
More precisely, for any pair $(s,t)$, there is a 
   single
      $(s,t|\Pred_{s,t})\in\mathcal{R}$  stating that $s$ and $t$ are related 
      if $\Pred_{s,t}$       is 
      \\True, i.e. the states are related when the variables in $s$ and $t$ verify the 
      predicate $\Pred_{s,t}$.
 FH-bisimulation is defined formally: 
 \begin{definition}[Strong FH-bisimulation]\label{def-FH-bisim} ~\\

@@

\begin{figure}[t]
   \includegraphics[width=5.5cm]{XFIG/SimpleProt-Schema}
   \includegraphics[width=6.5cm]{XFIG/SimpleProt2-Spec}
   \caption{Specification pNet Schema and Specification}
   \label{App:SimpleProt:Spec}

@@

\begin{figure}[t]
  \centerline{\includegraphics[width=12cm]{XFIG/SimpleProt2-pNet}}
  \caption{Composed pNet with the Simple Protocol Implementation}  \label{App:SimpleProt:Impl}
\end{figure}

@@

The corresponding generated Open Automaton was given in Figure \ref{SimpleProtCounter:SpecOA},
page \pageref{SimpleProtCounter:SpecOA}.

@@

\begin{lstlisting}[basicstyle=\scriptsize\ttfamily, language=java, frame=single]
SimpleProtImpl2
--------------------
SV6 SVs: <p_send(m), in(m), _>-->in(m)
SV7 SVs: <p_a, _, _>-->p_a, p_a!=(p_send(m))
SV8 SVs: <_, out(m, c), q_recv(m, c)>-->out(m, c)
SV9 SVs: <_, _, q_a>-->q_a, q_a!=(q_recv(m, c))
SV11 SVs: <_, _tau_, _>-->_tau_
Holes: P
Subnets: medium
- medium
- SV0 SVs: <in(m), _, _>-->in(m)
- SV1 SVs: <s_send(m, c), m_recv(m, c), _>-->_tau_
- SV2 SVs: <_, m_send(m), r_recv(m)>-->_tau_
- SV3 SVs: <s_ack, _, r_ack>-->_tau_
- SV4 SVs: <s_error, m_error, _>-->_tau_
- SV5 SVs: <_, _, out(m, c)>-->out(m, c)
- SV10 SVs: <_, _tau_, _>-->_tau_
- pLTS: Sender
- - Transition: s0---in(msg)-->s1, { s1_msg := msg s1_c := 0}
- - Transition: s1---s_send(s1_msg, s1_c)-->s2, {}
- - Transition: s2---s_ack-->s0, {}
- - Transition: s2---s_error-->s1, { s1_c := s2_c+1}
- - Variable: s1_c, State: s1, Value(s) = 0,s2_c+1,
- - Variable: s1_msg, State: s1, Value(s) = msg,
- pLTS: M1
- - Transition: a0---m_recv(msg, c)-->a1, { a1_msg := msg a1_c := c}
- - Transition: a1---m_send(a1_msg, a1_c)-->a0, {}
- - Transition: a1---_tau_-->a2, {}
- - Transition: a2---m_error-->a0, {}
- - Variable: a1_msg, State: a1, Value(s) = msg,
- - Variable: a1_c, State: a1, Value(s) = c,
- pLTS: Receiver
- - Transition: r0---r_recv(msg, c)-->r1, { r1_msg := msg r1_c := c}
- - Transition: r1---out(r1_msg, r1_c)-->r2, {}
- - Transition: r2---r_ack-->r0, {}
- - Variable: r1_msg, State: r1, Value(s) = msg,
- - Variable: r1_c, State: r1, Value(s) = c,
Holes: Q
\end{lstlisting}

Finaly, we list here, in raw format, the 12 OTs computed by the tool:

\begin{lstlisting}[basicstyle=\scriptsize\ttfamily, language=java, frame=single]
INFO  SimpleProtImpl2 - 
	{P|->p_send(m:sva_SV6:1:1)},
	[(msg=m:sva_SV6:1:1)/\((in(msg))=(in(m:sva_SV6:1:1)))],
	{s1_msg := msg, s1_c := 0}
OT:1 ------------------------------------------------------------
	<s0_a0_r0>-----in(m:sva_SV6:1:1)----><s1_a0_r0>

@@

INFO  SimpleProtImpl2 - 
	{},   [(s1_msg=m:sva_SV1:114:2)/\(s1_c=c:sva_SV1:114:2)
	  /\((s_send(s1_msg, s1_c))=(s_send(m:sva_SV1:114:2, c:sva_SV1:114:2)))
          /\(msg=m:sva_SV1:114:2)/\(c=c:sva_SV1:114:2)
          /\((m_recv(msg, c))=(m_recv(m:sva_SV1:114:2, c:sva_SV1:114:2)))],
	{a1_msg := msg, a1_c := c}
OT:6 ------------------------------------------------------------
	<s1_a0_r0>-----_tau_----><s2_a1_r0>

INFO  SimpleProtImpl2 - 
	{P|->p_a:sva_SV7:1:2},
	[(forall:m:sva_SV7:1:2. (p_a:sva_SV7:1:2!=(p_send(m:sva_SV7:1:2)))],   {}
OT:7 ------------------------------------------------------------
	<s2_a1_r0>-----p_a:sva_SV7:1:2----><s2_a1_r0>

INFO  SimpleProtImpl2 - 
	{Q|->q_a:sva_SV9:1:4},
	[(forall:m:sva_SV9:1:4,c:sva_SV9:1:4.
          (q_a:sva_SV9:1:4!=(q_recv(m:sva_SV9:1:4, c:sva_SV9:1:4)))],   {}
OT:8 ------------------------------------------------------------
	<s2_a1_r0>-----q_a:sva_SV9:1:4----><s2_a1_r0>

INFO  SimpleProtImpl2 - 
	{},   [],   {}
OT:9 ------------------------
	<s2_a1_r0>-----_tau_----><s2_a2_r0>

INFO  SimpleProtImpl2 - 
	{P|->p_a:sva_SV7:1:2},
	[(forall:m:sva_SV7:1:2. (p_a:sva_SV7:1:2!=(p_send(m:sva_SV7:1:2)))],   {}
OT:10 ------------------------------------------------------------
	<s2_a2_r0>-----p_a:sva_SV7:1:2----><s2_a2_r0>

INFO  SimpleProtImpl2 - 
	{Q|->q_a:sva_SV9:1:4},
	[(forall:m:sva_SV9:1:4,c:sva_SV9:1:4.
          (q_a:sva_SV9:1:4!=(q_recv(m:sva_SV9:1:4, c:sva_SV9:1:4)))],   {}
OT:11 ------------------------------------------------------------
	<s2_a2_r0>-----q_a:sva_SV9:1:4----><s2_a2_r0>

INFO  SimpleProtImpl2 - 
	{},   [],   {s1_c := s2_c+1}
OT:12 --------------------------------------
	<s2_a2_r0>-----_tau_----><s1_a0_r0>

INFO   SimpleProtImpl2 - State variables:

Global State: <s0_a0_r0>
===================
pLTS States: <s0 vars: {}> <a0 vars: {}> <r0 vars: {}>

Global State: <s1_a0_r0>
===================
pLTS States: <s1 vars: { s1_c <-{0, s2_c+1, } s1_msg <-{msg, }}> <a0 vars: {}> <r0 vars: {}>

Global State: <s2_a1_r0>
===================
pLTS States: <s2 vars: {}> <a1 vars: { a1_msg <-{msg, } a1_c <-{c, }}> <r0 vars: {}>

Global State: <s2_a2_r0>
===================
pLTS States: <s2 vars: {}> <a2 vars: {}> <r0 vars: {}>

INFO  SimpleProtImpl2 - Total: 12 SATISFIABLE OTs
INFO  SimpleProtImpl2 -        122 UNSATISFIABLE OTs
INFO  SimpleProtImpl2 - Total Time: 657 ms

@@
